The markings of the trade: exploring the patterns of olive oil production in Roman Baetica

### 2020-07-27 Paper/jaccardbaetica.tex
add new comments xavi almost

\xavi{reescribir abstract, demasiado general y entonces demasiado específico}
The aim of this study is to explore economics dynamics in the production and distribution of olive oil trade. 
Our case of study has been focused on the production processes located in \textit{Baetica} province (currently Andalusia) from 1st to 3rd AD. In particular, we want to detect patterns of olive oil production that link amphora workshops and amphoric stamps. \textit{Baetica} became an important production and distribution centre during the Roman Empire. However, it remains under debate about how this province was organised and whether it could be possible to identify patterns in the olive oil market. Amphoric stamps are used to identify the presence of different groups that might share similar stamps. To achieve this goal, we analyse a set of stamps from two centres: 1) production centres by analysing different workshops in \textit{Baetica} province and 2) two Roman provinces such as \textit{Germania} and \textit{Britannia} as consumption centres. They will be used to detect a connection between the distribution of amphoric stamps and the economic structure in both centres. Here, we use methods borrowed from Ecology that allow us to identify if amphora workshops share similar amphoric stamps depending on the spatial distance.

The analysis explores how the quantitative approach provides a useful tool for the interpretation of the economic processes. Finally, results pretend to highlight the organisation of Baetican olive oil production in the Roman Empire linked to the differences observed in the archaeological evidence.

@@

Most stamp codes display a large activity of production that can be difficult to date using an accurate chronology. They have been frequently dated with the consular dating by studying \textit{tituli picti} found in \textit{Monte Testaccio} \citep{Testaccio1, berni_millet_epigrafianforica_2008}. However, the chronology of stamps could be also biased in other sites taking into account that amphorae were deposited after being produced and consumed\xavi{esto necesita un poco de desarrollo. quieres decir que el código se encuentra en contextos de cronologías diferentes no? Y lo del Testaccio, pos eso, que hay pocos y todos de un yacimioento, lo que añade un bias}.

These stamps are typically found on handles but they were also imprinted on rims and the amphora body~\citep{millet_anforas_1998}. The content information of the stamps is shown in different forms and letter content and it seems that there was not a unique criterion.
%\xavi{qué information? El código de letras?}. 
Stamps often displayed a code of three letters and they are known as \textit{Tria Nomina} \citep{berni_millet_amphora_1996}. The three letters correspond to a initial forms of a name but they can also appear complete. 
%\xavi{que quieres decir con abbreviated o complete? Como abrevias un código de 3 letras?}

@@

The stamp codes are interpreted based on three main ideas: content (olive oil), context (amphora workshop) and subject (individuals involved on the production). On the one hand, it seems that stamps could have been identified as the landowner of the olive groves \citep{rodriguez_economioleicola_1977}. On the other hand, they could also belong to the owner of the workshops were the amphorae was made or even to a group of amphora workers~\citep{berni_millet_epigrafianforica_2008}. In any case, the use of these stamps is a good proxy to explore the system of olive oil production and trade developed in Baetica.

Nevertheless, some challenges\xavi{acabas de decir que tampoco se sabe qué es el sello, así que esta frase queda rara} remain under discussion such as how this production was organised and whether it is possible to distinguish production patterns in the olive oil trade. Our questions will be focused on the distribution of amphoric stamps. Did they follow a distribution pattern? Did stamps share the same workshop?

Neither the use of written records have allowed providing enough information that can explain the economic role of \textit{Baetica} province in the Roman organisation.\xavi{esta frase esta fuera de contexto. Si la pones entonces deberías expandir la idea de frase a párrafo, y poner referencias}.

@@

\xavi{faltaría un enlace a la siguiente sección, queda muy abrupto}

@@

It is important to mention that some workshops exhibited the same coordinates but they were catalogued as different workshops in Roman Studies (i.e. nearby amphora workshops in \textit{Baetica} named as Tesorillo de Doña Mencia and Doña Mencia). The analysis took into account this particularity but some biases can occur due to the archaeological dataset.

\xavi{No se entiende; deberías explicarlo un poco mejor y citar algo si dices eso. Piensa que la mayoría de lectores no sabrán quién es Doña Mencia}.

@@

This research was partly funded by the European Research Council Advanced Grant EPNet (340828). We are grateful to Simon Carrignon, Juan Moros and Ignacio Morer for their useful suggestions.  
All data has been analysed and conducted in R program version 3.2.4, using the packages \textit{vegan} \citep{oksanen_vegan_2007}, \textit{ggplot2} \citep{ggplot2:_2016}. Source and code are available at \maria{incluir github}.

>>>>

\xavi{reescribir abstract, demasiado general y entonces demasiado específico}
\maria{lo he reeformulado a ver si ahora..}

The aim of this study is to explore economics dynamics in the production and distribution of olive oil trade. 
Baetica became an important production and distribution centre during the Roman Empire. However, it remains under debate about how this province was organised and whether it could be possible to identify patterns in the olive oil market.

Our case of study has been focused on the production processes located in Baetica province (currently Andalusia) from 1st to 3rd AD. In particular, we want to detect patterns of olive oil production that link amphora workshops and amphoric stamps. Amphoric stamps are used to identify the presence of different groups that might share similar stamps. To achieve this goal, we analyse a set of stamps from two centres: 1) production centres by analysing different workshops in Baetica province and 2) two Roman provinces such as \textit{Germania} (Inferior and Superior) and \textit{Britannia} as consumption centres. They will be used to detect a connection between the distribution of amphoric stamps and the economic structure in both centres. Here, we use methods borrowed from Ecology that allow us to identify if amphora workshops share similar amphoric stamps depending on the spatial distance. The analysis explores how the quantitative approach provides a useful tool for the interpretation of the economic processes. Finally, results pretend to highlight the organization of Baetican olive oil production in the Roman Empire linked to the differences observed in the archaeological evidence.

%Our case of study has been focused on the production processes located in \textit{Baetica} province (currently Andalusia) from 1st to 3rd AD. In particular, we want to detect patterns of olive oil production that link amphora workshops and amphoric stamps. \textit{Baetica} became an important production and distribution centre during the Roman Empire. However, it remains under debate about how this province was organised and whether it could be possible to identify patterns in the olive oil market. Amphoric stamps are used to identify the presence of different groups that might share similar stamps. To achieve this goal, we analyse a set of stamps from two centres: 1) production centres by analysing different workshops in \textit{Baetica} province and 2) two Roman provinces such as \textit{Germania} and \textit{Britannia} as consumption centres. They will be used to detect a connection between the distribution of amphoric stamps and the economic structure in both centres. Here, we use methods borrowed from Ecology that allow us to identify if amphora workshops share similar amphoric stamps depending on the spatial distance. The analysis explores how the quantitative approach provides a useful tool for the interpretation of the economic processes. Finally, results pretend to highlight the organisation of Baetican olive oil production in the Roman Empire linked to the differences observed in the archaeological evidence.

@@

Most stamp codes display a large activity of production that can be difficult to date using an accurate chronology. They have been frequently dated with the consular dating by studying \textit{tituli picti} found in \textit{Monte Testaccio} \citep{Testaccio1, berni_millet_epigrafianforica_2008}.
Despite \textit{tituli picti} contain important information about chronology, they have been still found in few quantities mostly concentrated in a only settlement such as \textit{Monte Testaccio}. In addition, the chronology of stamps could be also biased taking into account that amphorae were deposited after being produced and consumed\xavi{esto necesita un poco de desarrollo. quieres decir que el código se encuentra en contextos de cronologías diferentes no? Y lo del Testaccio, pos eso, que hay pocos y todos de un yacimioento, lo que añade un bias}. 
\maria{desarrollado}

The use of amphoric stamps can be a good proxy to explore the system of olive oil production and trade developed in Baetica province. These stamps are typically found on handles but they were also imprinted on rims and the amphora body~\citep{millet_anforas_1998}. The content information of the stamps is shown in different forms and letter content and it seems that there was not a unique criterion.
%\xavi{qué information? El código de letras?}. 
Stamps often displayed a code of three letters and they are known as \textit{Tria Nomina} \citep{berni_millet_amphora_1996}. The three letters correspond to a initial forms of a name but they can also appear complete. 
%\xavi{que quieres decir con abbreviated o complete? Como abrevias un código de 3 letras?}

@@

The stamp codes are interpreted based on three main ideas: content (olive oil), context (amphora workshop) and subject (individuals involved on the production). On the one hand, it seems that stamps could have been identified as the landowner of the olive groves \citep{rodriguez_economioleicola_1977}. On the other hand, they could also belong to the owner of the workshops were the amphorae was made or even to a group of amphora workers~\citep{berni_millet_epigrafianforica_2008}. 
%In any case, the use of these stamps is a good proxy to explore the system of olive oil production and trade developed in Baetica.

\xavi{acabas de decir que tampoco se sabe qué es el sello, así que esta frase queda rara} 
\maria{he cambiado la frase de sitio para que se contextualice mejor}
Nevertheless, some challenges remain under discussion such as how this production was organised and whether it is possible to distinguish production patterns in the olive oil trade. Our questions will be focused on the distribution of amphoric stamps. Did they follow a distribution pattern? Did stamps share the same workshop?

%Neither the use of written records have allowed providing enough information that can explain the economic role of \textit{Baetica} province in the Roman organisation.\xavi{esta frase esta fuera de contexto. Si la pones entonces deberías expandir la idea de frase a párrafo, y poner referencias}.
\maria{he eliminado la última frase}

@@

The Roman conquest in \textit{Britannia} and \textit{Germania} was a new opening to foreign trade in order to supply the military campaign. A high demand for olive oil in both provinces stimulated an important trade network involving different agents and the Empire itself. Baetica became a strategic province for the production and distribution in both provinces since it can be detected by the high presence of amphoric stamps found in civil and military areas during this period.

\xavi{faltaría un enlace a la siguiente sección, queda muy abrupto}
\maria{he añadido varios párrafos a ver si se enlaza}

@@

It is important to mention that some workshops exhibited the same geographical coordinates while they were catalogued differently as workshops. For instance, two nearby amphora workshops such as Tesorillo de Doña Mencia and Doña Mencia were catalogued as a different name in spite of having the same geographical coordinates.  Our work considered this particularity in the archaeological dataset but some biases can occur in the analysis.

%It is important to mention that some workshops exhibited the same coordinates but they were catalogued as different workshops in Roman Studies (i.e. nearby amphora workshops in \textit{Baetica} named as Tesorillo de Doña Mencia and Doña Mencia). The analysis took into account this particularity but some biases can occur due to the archaeological dataset.

\xavi{No se entiende; deberías explicarlo un poco mejor y citar algo si dices eso. Piensa que la mayoría de lectores no sabrán quién es Doña Mencia}. 
\maria{he reformulado la frase a ver si ahora se entiende}

@@

This research was partly funded by the European Research Council Advanced Grant EPNet (340828). We are grateful to Simon Carrignon, Juan Moros, Víctor M. Lozano and Ignacio Morer for their useful suggestions.  
All data has been analysed and conducted in R program version 3.2.4, using the packages \textit{vegan} \citep{oksanen_vegan_2007}, \textit{ggplot2} \citep{ggplot2:_2016}. Source and code are available at \maria{incluir github}.
